MADA Project_UOWN

### 2021-09-17 UOWN_MADA_CC.Rmd
Project Part 1 - CC

rawdata <- readxl::read_excel(data_location)
```

@@

#When monitoring freshwater streams, stream health is typically measured by its physical charachteristics (chemistry, surrounding land use area, channelication, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates). However, these thre categories are usually analyzed separately to assess stream health. As such, my question is how well are these parameters correleated with each other. Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?

I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in on indicator category will indicate a correlated negative impact in the other categories. The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find.

@@

UOWN has provided a great deal of data dating back to 2003. First,I will analyze and compare trends in the change of chemical (conductivity, turbidity,	po4.mgL,	no3.mgL); ecological(Biological Score); and microbial indicator data (e.coli CFU). Generally speaking, Increases in chemical properties like conductivity, turbidity,	po4.mgL,	no3.mgL and microbial indicator abundance (i.e. ecoli CFU) indicate decreased stream health. While a decrease in biological score indicates decrease in stream health. Next, since biological scores are used as a measure of broader ecological health, I will run t-tests on if the other factors significantly effect biological score. Finally, I will test the interaction of all three vie multivariate analysis.

>>>>

rawdata <- readxl::read_excel(data_location)

summary(rawdata)
```

@@

When monitoring freshwater streams, stream health is typically measured by its physical charachteristics (chemistry, surrounding land use area, channelication, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates). However, these thre categories are usually analyzed separately to assess stream health. As such, my question is how well are these parameters correleated with each other. Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?

I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact in the other categories. The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find.

@@

UOWN has provided a great deal of data dating back to 2003. First,I will need to wrangle missing data values that are a result of inconsitent sampling efforts. This will make the data easier to work with and narrow down a time frame for the data that is consistent and free of sampling bias. Next, I will analyze and compare trends in the change of chemical (conductivity, turbidity,	po4.mgL,	no3.mgL); ecological(Biological Score); and microbial indicator data (e.coli CFU). Generally speaking, Increases in chemical properties like conductivity, turbidity,	po4.mgL,	no3.mgL and microbial indicator abundance (i.e. ecoli CFU) indicate decreased stream health. While a decrease in biological score indicates decrease in stream health. Next, since biological scores are used as a measure of broader ecological health, I will run t-tests on if the other factors significantly effect biological score. Finally, I will test the interaction of all three vie multivariate analysis.
